Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-02-01 main.tex
Writing down the stack safety formalization based on yesterdays talk

@@

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\VIS,\HID\}$ is a map
from machine components to ``observabilities,'' where
$\VIS$ marks locations whose contents are observable from the outside (by
the ``attacker,'' e.g., by DMA) and $\HID$ marks internal ``hidden''
state that cannot be observed directly (e.g., processor registers and memory
addresses for temporary variables will usually be hidden).

A {\em contoured trace} of length $n$, where $n$ can be finite or
infinite, is a length-$n$ sequence $\seq{(\contour,\mach)}_n =
(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$ of (contour,
machine state) pairs in which adjacent machine states are related by
$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
\mach_i$ for each $i > 0$ in the sequence.

\subsection{Observations}

A {\em component observation} is either a word or the special value $\bot$.
A {\em state observation} is a map from machine components to component
observations.

An {\em observation} is a nonempty, finite or infinite sequence of state
observations in which (1) adjacent elements are always distinct (``no
stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
the the $j^{th}$ state observation in the sequence must also be mapped to
$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
component's value changes from $\bot$ to anything else, it never changes
back to $\bot$ again).

In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
to its value in $\mach_0$ and every hidden
component to $\bot$; 
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
visible component (where $\contour_{i+1} = \VIS$), copying its
value from $\mach_{i+1}$, and (2) for each
hidden component, where $\contour_i = \VIS$, copying its
value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
\iftext\bcp{Too informal!!}\fi
\end{itemize}

@@

\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

A {\em contoured hyperproperty} is a set of pairs of contoured
traces.

A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
such that:

\leo{Informally first, I don't like it too much:}
\begin{itemize}
\item each trace in every pair of a contoured trace in the set have the
  same length, and
\item each pair of elements in every pair of a contoured trace in the
  set have the same contour.
\end{itemize}

\leo{Formally:}
\begin{itemize}
\item {\em same-length:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
    M')}_m$ in the set, $n = m$.

\item {\em contour-lockstep:}

  For every pair of traces $\overline{(C, M)}_n$ and $\overline{(C',
    M')}_n$ in the set, $C_i = C_i'$ for all $i<n$.
\end{itemize}

Finally, we can parameterize the traces by some initial
application-specific state to obtain a {\em contoured lockstep
  $X$-hyperproperty}, where every trace comes with more information.
\leo{Example, call sites in stack safety}

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\VIS$ by $\contour$---i.e., if $\contour(\component) = \VIS$, then
$\mach(\component) = \mach'(\component)$.
\apt{Um..this seems to conflate visibility (\VIS,\HID) with
  security/privacy level (\HIGH,\LOW).  Don't we want to distinguish these?}

\leo{Normal lockstep noninterference:}

\leo{For this the X- component is empty right?}
Lockstep-Noninterference is the contoured lockstep hyperproperty
where:
\begin{itemize}
\item For each pair of contoured traces in the hyperproperty, the
  contour is invariant along each trace.
\item For each pair of contoured traces in the hyperproperty, if the
  initial elements of the traces are variants, then the observervations
  of the two traces coincide.
\end{itemize}

\leo{Nested lockstep noninterference:}

Lockstep-Noninterference is a restricted form of traditional
noninterference, that is nevertheless useful (e.g. to encode taint
tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
At the same time, lockstepness makes it possible to generalize the
setting and allow for the contours themselves to change throughout the
execution.

@@

>>>>

\begin{itemize}
\item \leo{Not even sure this is true anymore}
\item Noninterference is a security property combining {\em
  confidentiality} and {\em integrity}: secret data should influence
  public outputs and low integrity data should \leo{finish in a pretty
    way}.
\item In this paper we present a stack safety property as a variation
  of noninterference.
\end{itemize}

@@

\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGH,\LOW\} \times \{\HI, \LO\}$ is a map from machine components
to confidentiality and integrity levels. $\LOW$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGH$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LO$
locations correspond to low-integrity components (e.g. locations that
can be written to by the network), while $\HI$ locations should be
protected from such flows.

% 
%\subsection{Observations}
% 
%A {\em component observation} is either a word or the special value $\bot$.
%A {\em state observation} is a map from machine components to component
%observations.
% 
%An {\em observation} is a nonempty, finite or infinite sequence of state
%observations in which (1) adjacent elements are always distinct (``no
%stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
%the the $j^{th}$ state observation in the sequence must also be mapped to
%$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
%component's value changes from $\bot$ to anything else, it never changes
%back to $\bot$ again).
% 
%In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
%from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
%observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
%follows:
%\begin{itemize}
%\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
%to its value in $\mach_0$ and every hidden
%component to $\bot$; 
%\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
%visible component (where $\contour_{i+1} = \VIS$), copying its
%value from $\mach_{i+1}$, and (2) for each
%hidden component, where $\contour_i = \VIS$, copying its
%value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
%identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
%\iftext\bcp{Too informal!!}\fi
%\end{itemize}

@@

%\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

A {\em trace} of length $n$, where $n$ can be finite or infinite, is a
length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
machine states in which adjacent machine states are related by
$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
\mach_i$ for each $i > 0$ in the sequence.

A {\em contoured hyperproperty} is a set of triples: a contour along
with a pair of traces.

\leo{No longer needed:?}
A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
such that each trace in every pair of traces in the set executes the
same instruction sequence.

Finally, we can parameterize the traces by some initial
application-specific state to obtain a {\em contoured lockstep
  $X$-hyperproperty}, where every trace comes with more information,
for example call sites in stack safety.

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\LOW$ by $\contour$---i.e., if $\contour(\component) = \LOW$, then
$\mach(\component) = \mach'(\component)$.

\leo{Baseline noninterference:}

\leo{Is this noninterference or is it noninterference + integrity.
  I'm confused about terminology}

Our baseline security property is noninterference.
\begin{itemize}
\item {\em Confidentiality:}

  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$, if
  the initial elements of the traces are variants, then the final
  states of the traces are also variants.

\item {\em Integrity: }

  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$,
  and for every high-integrity component $\component$ for which
  $\contour(\component) = \HI$, $M_0(\component) = M_n(\component)$
  and $M'_0(\component) = M'_{n'}(\component)$.
\end{itemize}
\leo{In the terminology of the original hyperproperty paper,
  confidentiality is a hyperproperty, integrity is a trace property
  that holds for every trace in a hyperproperty. This is a stronger
  version of integrity than DLM for example}.
% 
%\leo{Nested lockstep noninterference:}
% 
%Lockstep-Noninterference is a restricted form of traditional
%noninterference, that is nevertheless useful (e.g. to encode taint
%tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
%At the same time, lockstepness makes it possible to generalize the
%setting and allow for the contours themselves to change throughout the
%execution.

@@

\leo{To warm up, a super simple separation policy.}
\begin{itemize}
\item Any instruction can serve as a final instruction.\leo{What about
  instructions the are halted by the monitor?}
\item For forward compatibility we could treat the first unmatched return as
  a halt instruction.
\item The instruction memory is marked $\LOW \times \HI$. That means
  it is observable and can't change.
\item The memory space is split in two. One part is marked $\LOW
  \times \LO$: this is the public part (or the one that can be
  accessed by the current process e.g.). The other part is marked
  $\HIGH \times \HI$: it should be completely inaccessible to the
  current process.
\item Separation is noninterference:
  \leo{Exactly the definition above}
\item Depending on what we do with final instructions, the property
  could be parameterized by the ``call/return sites''.
\end{itemize}
\leo{(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step because of integrity?)}

\section{Stack Safety is noninterference$+$}

\leo{Build up to the interesting thing:}
\begin{itemize}
\item The first unmatched return serves as a halt.
\item The instruction memory is again marked $\LOW \times \HI$. That
  means it is observable (both traces have the same instructions!) and
  can't change.
\item At the beginning of time, the entire data memory is marked $\LOW
  \times \LO$. That makes for a very unintersting property as all
  variants are necessarily identical.
\item BUT, we close the hyperproperty as follows:
  \begin{itemize}
  \item We need the state description stuff to enforce WBCF and be
    able to talk about ``first unmatched returns''. That is also a
    trace property.
  \item After every call, we keep the part of the (left-hand, wlog
    due to symmetry) trace until its first unmatched return.
  \item We construct a new contour at that point: after every call,
    the local state of the caller becomes inaccessible (marked
    $\HIGH \times \HI$) until a subsequent return.
  \item For every variant of the new initial state with respect to the
    new contour, the contour plus pair of traces is in the
    hyperproperty.
  \end{itemize}  
\end{itemize}

\section{Old Stuff}
